---
title: "SubwatershedAssessment"
format: html
editor: visual
---

# Import watershed shape files

The following code chunks follow a similar analysis pattern as described here.

1.  Imports watershed shapefile
2.  Extracts modeled land cover by watershed outline
3.  Segregates each land cover type so spatial patterns can be more easily observed
4.  Calculates area of total watershed and each land cover type
5.  Extracts elevation based on watershed boundaries

### Tundra sub-watershed

```{r}
#upload shapefiles, set crs
tundra.shed<- st_read("/Users/hannahrichardson/Documents/BSU/Research/Data/Landsat Data/Watershed boundaries/CRS works/TundraSubwatershed")
tundra.shed<- terra::vect(st_transform(tundra.shed, crs = st_crs(landsat.st)))
prlandcov<-terra::rast(pr21)
tundra.shed<-project(tundra.shed, prlandcov)
tundra.cover<- terra::crop(prlandcov,tundra.shed, mask=TRUE)
tundra.shed.size<-terra::expanse(tundra.cover, unit="km")

#segment by classification
tundra.segments<-terra::segregate(tundra.cover)
names(tundra.segments) <- c("forest","ice","rockwater","shrub","tundra")
plot(tundra.segments)

#calculate area size for each cover type
forest.size<-sum(values(tundra.segments$forest, na.rm=TRUE))
ice.size<-sum(values(tundra.segments$ice), na.rm=TRUE)
rockwater.size<-sum(values(tundra.segments$rockwater), na.rm=TRUE)
shrub.size<-sum(values(tundra.segments$shrub), na.rm=TRUE)
tundra.size<-sum(values(tundra.segments$tundra), na.rm=TRUE)

#calculate total area and percents
tundra.all.size<- sum(forest.size, ice.size, rockwater.size, shrub.size, tundra.size)
forper<-(forest.size/tundra.all.size)*100
iceper<-(ice.size/tundra.all.size)*100
rowaper<-(rockwater.size/tundra.all.size)*100
shrper<-(shrub.size/tundra.all.size)*100
tunper<-(tundra.size/tundra.all.size)*100
tundra.percents<-c(forper, iceper, rowaper, shrper, tunper)

tundra.percents
AKtundra<-terra::project(AK,crs(tundra.shed)) 
tundra.elev<-terra::crop(AKtundra, tundra.shed, mask=TRUE)
```

### Forest sub-watershed

```{r}
#upload shapefiles, set crs
forest.shed<- st_read("/Users/hannahrichardson/Documents/BSU/Research/Data/Landsat Data/Watershed boundaries/CRS works/Forest_New")
forest.shed<- terra::vect(st_transform(forest.shed, crs = st_crs(landsat.st)))
prlandcov<-terra::rast(pr21)
forest.shed<-project(forest.shed, prlandcov)
forest.cover<- terra::crop(prlandcov,forest.shed, mask=TRUE)
forest.shed.size<-terra::expanse(forest.cover, unit="km")

#segment by classification
forest.segments<-terra::segregate(forest.cover)
names(forest.segments) <- c("forest","rockwater")
plot(forest.segments)

#calculate area size for each cover type
#some subwatersheds won't have every type of land cover, I have been manually commenting those out for this step

forest.size<-sum(values(forest.segments$forest, na.rm=TRUE))
#ice.size<-sum(values(forest.segments$ice), na.rm=TRUE)  
rockwater.size<-sum(values(forest.segments$rockwater), na.rm=TRUE)
#shrub.size<-sum(values(forest.segments$shrub), na.rm=TRUE)
#tundra.size<-sum(values(forest.segments$tundra), na.rm=TRUE)

#calculate total area and percents
all.size<- sum(forest.size, rockwater.size)
forper<-(forest.size/all.size)*100
#iceper<-(ice.size/all.size)*100
rowaper<-(rockwater.size/all.size)*100
#shrper<-(shrub.size/all.size)*100
#tunper<-(tundra.size/all.size)*100
forest.percents<-c(forper, iceper, rowaper, shrper, tunper) 

forest.percents
#forest.elev<-terra::crop(AKnj, forest.shed, mask=TRUE)
```
